---
title: "Assignment 1"
author: "Sagardweep Saha- 19200237"
date: "11/10/2019"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
install.packages('tinytex')
## Load the Dataset Eurocrime
```{r ,Read}
Eurocrime<-read.csv("C:/Autumn/Data Programming with R/Assignment/EurostatCrime2017.csv",
header=TRUE,row.names=1)
```

## Display the number of obervations and variables in the dataframe
```{r, Number of row and column}
nrow(Eurocrime)
ncol(Eurocrime)
dim(Eurocrime)
```
Ans: 41 Observations and 11 Variables

## Adding a new column called ALL_Theft
```{r, Add a new column called All_Theft}
Eurocrime["ALL_Theft"]<- 0
Eurocrime$ALL_Theft<-rowSums(Eurocrime[,c(7,8,9,10)],na.rm = TRUE)
```

## Removing a few columns from the original dataframe and creating a new data frame
```{r, remove a few columns}
Eurocrime_new<- Eurocrime[-c(7,8,9,10)]
```

## Displaying the countries with missing data
```{r, list counntries with missing data}
row.names(Eurocrime_new[!complete.cases(Eurocrime_new),])
```
Ans:  Denmark,France,Croatia,Hungary,
Netherlands,Austria,Poland,Portugal
Sweden,England_and_Wales,Iceland,
Liechtenstein,Norway,North_Macedonia,Turkey,Bosnia_and_Herzegovina
 
## Removing the countries with missing data and creating a new dataframe
```{r, remove the countries with missing data}
Eurocrime_Refined<-na.omit(Eurocrime_new)
Eurocrime_Refined
```

## Displaying the number of observations and variables in Eurocrime_Refined
```{r, number of observations and variables in the dataframe}
nrow(Eurocrime_Refined)
ncol(Eurocrime_Refined)
dim(Eurocrime_Refined)
```
Ans: 25 observations and 8 Variables

## Displaying the 3 most common crimes in Ireland in 2017
```{r, three most common crimes in Ireland in 2017}
row=which(rownames(Eurocrime_Refined)=="Ireland")
ireland<-Eurocrime_Refined[row,]
Ireland_sorted<-sort(ireland, decreasing  = TRUE)
Ireland_sorted[,1:3]

sort(Eurocrime_Refined[c("Ireland"),],decreasing = T)[1:3]
```
Ans: The three most common crimes in Ireland in 2017 are:
All_Theft=1956.29, Unlawful_acts_involving_
controlled_drugs_or_precursors= 351.58, Assault=84.59

## Displaying the country that has the highest overall record of offences (per hundred thousand inhabitants)
```{r, country has the highest overall record of offences}
Eurocrime_Refined["Total_Offences"]<-0
Eurocrime_Refined$Total_Offences<-rowSums(Eurocrime_Refined, na.rm = TRUE)
max(Eurocrime_Refined$Total_Offences, na.rm=TRUE)
Eurocrime_Refined[match(3515.89, Eurocrime_Refined$Total_Offences),]
```
Ans: Switzerland

## Plotting two barplots where we analyze the data between two columns 
## i.e sexual_violence v/s Eurocrime_Refined$Unlawful_acts_involving_controlle
## d_drugs_or_precursors. Here we can see that in most countries 
## where acts realted to drugs are more, sex violence is quite low.
```{r, Analysis}

h<-cbind(Eurocrime_Refined$Sexual.violence,
Eurocrime_Refined$Unlawful_acts_involving_controlled_drugs_or_precursors)
barplot(h, beside = TRUE, xlab = "sexual_violence v/s Eurocrime_Refined$Unlawful_acts_invo
lving_controlled_drugs_or_precursors ")
```